---
title: "PA1_template"
author: "Lucas Horiye"
date: "Tuesday, September 15, 2015"
output: html_document
---

# Assignment

## Loading and preprocessing data

Loading package reading file and transforming data

```{r results="hide", message=FALSE, warning=FALSE}
library(dplyr)
```

```{r echo=TRUE}
file <- read.csv('activity.csv')
activity_tbl <- tbl_df(file)
```

## What is mean total number of steps taken per day?

Total number of steps taken per day
```{r echo=TRUE}
steps_day <- activity_tbl %>% group_by(date) %>% summarize(steps_sum =sum(steps, na.rm=TRUE))
```
Histogram of the total number of steps taken each day
```{r echo=TRUE}
hist(steps_day$steps_sum, xlab="total steps per day", main="Histogram of total steps per day")
```

Mean and Median
```{r echo=TRUE}
mean(steps_day$steps_sum)
median(steps_day$steps_sum)
```

## What is the average daily activity pattern?

Overall plot
```{r echo=TRUE}
steps_interval_mean <- activity_tbl %>% group_by(interval) %>% summarize(steps_interval_mean =mean(steps, na.rm=TRUE))
with(steps_interval_mean, plot(interval, steps_interval_mean, type="l"))
```

A little bit closer (the answer is 835)
```{r echo=TRUE}
with(steps_interval_mean, plot(interval, steps_interval_mean, type="l", xlim=c(832, 838)))
```

## Imputing missing values

Total number of missing values
```{r echo=TRUE}
nrow(activity_tbl %>% filter(is.na(steps)))
```

Filling dataset
(STRATEGY: Using mean for that 5-minute-interval)
```{r echo=TRUE}
activity_tbl_filled <- activity_tbl %>% inner_join(steps_interval_mean)
activity_tbl_filled <- activity_tbl_filled %>% mutate(steps = ifelse(is.na(steps), steps_interval_mean, steps))
```

Calculating Mean and Median
```{r echo=TRUE}
steps_day_filled <- activity_tbl_filled %>% group_by(date) %>% summarize(steps_sum =sum(steps))

mean(steps_day_filled$steps_sum)
median(steps_day_filled$steps_sum)

```

## Are there differences in activity patterns between weekdays and weekends?

```{r results="hide", message=FALSE, warning=FALSE}
Sys.setlocale("LC_ALL","English")
```

```{r echo=TRUE}
activity_tbl_filled <- mutate(activity_tbl_filled, steps_interval_mean=NULL)
activity_tbl_filled <- mutate( activity_tbl_filled, wday =  ifelse 
( grepl ('Saturday|Sunday', weekdays(as.Date(date)) ), 'weekend', 'weekday'))
steps_wday <- activity_tbl_filled %>% group_by(interval, wday) %>% summarize(steps_mean = mean(steps))
```
```{r results="hide", message=FALSE, warning=FALSE}
library(lattice)
```
```{r echo=TRUE}
xyplot(data=steps_wday, steps_mean ~ interval | wday, type="l", layout=c(1, 2), main="", ylab="Number of steps", xlab="Interval")
```
